---
title: "Forests of the Future: Week 6 Assignment"
output: 
  learnr::tutorial:
    progressive: false
    language:
      en:
        button:
          runcode: Test Code
          submitanswer: Run Code
          questionsubmit: Submit Answer
          questiontryagain: Change my answer
runtime: shiny_prerendered
description: Forests of the Future Week 6 Assignment
---

```{css echo=FALSE}
@media print {
  .topicsContainer,
  .topicActions,
  .exerciseActions .skip {
    display: none;
  }
  .topics .tutorialTitle,
  .topics .section.level2,
  .topics .section.level3:not(.hide) {
    display: block;
  }
  .topics {
    width: 100%;
  }
  .tutorial-exercise, .tutorial-question {
    page-break-inside: avoid;
  }
  .section.level3.done h3 {
    padding-left: 0;
    background-image: none;
  }
  .topics .showSkip .exerciseActions::before {
    content: "Topic not yet completed...";
    font-style: italic;
  }
}
```

```{r setup_hide, include=FALSE}
library(learnr)
library(fofpack)

custom_checker <- function(label, user_code, check_code, envir_result, evaluate_result, envir_prep, last_value, stage, ...) {
  # this is a code check
  if(stage == "check") {
    
    fofpack::send_env_to_RStudio(envir_prep)
    
    rstudioapi::sendToConsole(user_code, focus = TRUE)
    
    fofpack::set_env(envir_result)
    
    try(save.image(file.path(system.file(package = "fofpack"), "week_5_progress.Rdata"),
               safe = FALSE),
        silent = TRUE)

    list(message = "Code Run; Results Now Available in RStudio.", correct = TRUE, type = "success", location = "append")
    
  }

}

tutorial_options(exercise.checker = custom_checker,
                 exercise.timelimit = 300)

knitr::opts_chunk$set(echo = FALSE)
```

## Forests of the Future Week 6 Assignment

### Forida Pine Rocklands

In this assignment we will be working with  Pine Rockland 

### Setup

```{r setup, include=FALSE}
library(fofpack)
library(tidyverse)
library(tidymodels)

data("RF_abund")

progress <- file.path(system.file(package = "fofpack"), "week_5_progress.Rdata")
if(file.exists(progress)) {
  load(progress, envir = rlang::global_env())
}
```

The following code loads the packages you will need for this assignment, and the dataset you will be using (and clears the workspace). You will be using a dataset on reef fishes from reefs around the world called `RF_abund`. Don't forget to always click 'Run Code' for each code box, which runs your code in RStudio, and allows you to access the results in later code boxes.

```{r do_setup, exercise=TRUE}
rm(list = ls())
library(fofpack)
library(tidyverse)
library(tidymodels)

data("RF_abund")
```

```{r do_setup-check}
ls()
```