---
title: 'Forests of the Future: Week 10 Lecture Part 2'
output:
  prettydoc::html_pretty:
    theme: tactile
    highlight: github
    self_contained: yes
description: Forests of the Future Week 10 Lecture Part 2
---

```{css echo = FALSE}
@media print {
  .topicsContainer,
  .topicActions,
  .exerciseActions .skip {
    display: none;
  }
  .topics .tutorialTitle,
  .topics .section.level2,
  .topics .section.level3:not(.hide) {
    display: block;
  }
  .topics {
    width: 100%;
  }
  .tutorial-exercise, .tutorial-question {
    page-break-inside: avoid;
  }
  .section.level3.done h3 {
    padding-left: 0;
    background-image: none;
  }
  .topics .showSkip .exerciseActions::before {
    content: "Topic not yet completed...";
    font-style: italic;
  }
}

```

```{r setup, include = FALSE}
library(ENMTools)
library(tidyverse)
library(tidymodels)
library(spatialsample)
library(tidysdm)
library(sf)
library(mapview)

options(datatable.alloccol = 1024,
        datatable.verbose = FALSE)
## skip rendering for now
knitr::opts_chunk$set(eval = FALSE)
```

## Last class

- We saw how to add 'pseudo-absences' to a presence-only data set to make modelling possible


## This class

- How to do spatial cross validation on presence / pseudo-absence data
- Talk about the final project



## An example SDM

- Let's run through a full SDM analysis
- We will use an example dataset in the `ENMTools` package


```{r get_points}
library(ENMTools)
library(tidyverse)
library(tidymodels)
library(spatialsample)
library(tidysdm)
library(sf)
library(mapview)

data("iberolacerta.clade")
data("euro.worldclim")
monticola <- iberolacerta.clade$species$monticola

monticola <- st_as_sf(monticola$presence.points, 
                      coords = c("Longitude", "Latitude"),
                      crs = 4326)
mapview(monticola)
```

## Create Background Area

```{r bg}
bg <- create_background(monticola, method = "ecoregion")
mapview(bg)

```

## Sample Pseudo-Absences

```{r bg_pnts}
monticola_dat <- sdm_data(monticola, bg = bg, n = 10000)
monticola_dat

mapview(monticola_dat, zcol = "present")

```

## Add environmental variables

```{r env}
monticola_dat <- add_env_vars(monticola_dat, euro.worldclim)
monticola_dat
```

## Remove NAs

```{r process}
monticola_dat <- monticola_dat %>%
  drop_na(bio1:bio19)

mapview(st_sf(monticola_dat), zcol = "bio1")
```

## Spatial Cross Validation

```{r cv}
## presence only (po) spatial CV
cv_folds_spat <- po_spatial_buffer_vfold_cv(monticola_dat, presence = "present", n = c(24, 16),
                                             v = 9)

## look at the spatial folds
autoplot(cv_folds_spat)

autoplot(cv_folds_spat$splits[[2]])
autoplot(cv_folds_spat$splits[[8]])

## regular CV for comparison
cv_folds <- vfold_cv(monticola_dat, 9)
```

## Make a recipe

- Let's do some transformations


```{r recipe}
monticola_recipe <- recipe(monticola_dat,
                         vars = c("present",
                                  as_tibble(monticola_dat) %>% select(bio1:bio19) %>% colnames()),
                         roles = c("outcome", rep("predictor", 19))) %>%
  step_YeoJohnson(all_predictors()) %>%
  step_normalize(all_predictors())

monticola_recipe
```

## Make Model and Workflow

```{r wf}
monticola_mod <- rand_forest() %>%
  set_engine("ranger", importance = "impurity") %>%
  set_mode("classification")

monticola_wf <- workflow() %>%
  add_recipe(monticola_recipe) %>%
  add_model(monticola_mod)

monticola_wf
```

## Fit Workflow

```{r fit}
monticola_fit <- monticola_wf %>%
  fit_resamples(cv_folds,
                control = control_resamples(extract = extract_fit_engine))

monticola_fit %>%
  collect_metrics()
```

## Fit Workflow with Spatial CV

```{r fit_spat}
monticola_fit_spat <- monticola_wf %>%
  fit_resamples(cv_folds_spat,
                control = control_resamples(extract = extract_fit_engine))

monticola_fit_spat %>%
  collect_metrics()
```

## Look at folds separately

```{r metrics_folds}
monticola_fit_spat$.metrics

```

## Importance

Let's have a look at the importance values determined by the random forest for our variables.


```{r importance}
library(vip)
library(patchwork)

monticola_fit %>%
  unnest(.extracts) %>%
  pull(.extracts) %>%
  map(vip) %>%
  wrap_plots(ncol = 3, nrow = 3)

```

The ordering is reasonably consistent between different folds. Now, the spatial folds:

---


```{r importance2}
monticola_fit_spat %>%
  unnest(.extracts) %>%
  pull(.extracts) %>%
  map(vip) %>%
  wrap_plots(ncol = 3, nrow = 3)

```

## Make a prediction dataset

- We want to make predictions on a landscape
- Use a grid:


```{r grid}
monticola_grid <- sdm_data(monticola, bg,
                         5000, sample_options = list(type = "regular")) %>%
  filter(present == "absent")

mapview(monticola_grid)
```

## Add environmental variable and predict

```{r grid_env}
monticola_grid_dat <- add_env_vars(monticola_grid, euro.worldclim) %>%
  drop_na(bio1:bio19)

monticola_grid_dat

final_fit <- monticola_wf %>%
  fit(monticola_dat)

monticola_preds <- final_fit %>% 
  augment(monticola_grid_dat)
```

## Plot predictions

```{r plot_pred}
coords <- st_coordinates(st_sf(monticola_preds))
ggplot(monticola_preds %>% bind_cols(coords), aes(X, Y)) +
  geom_sf(data = bg, inherit.aes = FALSE) +
  geom_raster(aes(fill = .pred_present + 0.0001)) +
  geom_sf(data = monticola, inherit.aes = FALSE, colour = "red",
          size = 1.2) +
  scale_fill_continuous(name = "Probability of Occurrence",
                        trans = "logit") +
  theme_minimal()
```

## Final Projects

- Next week we will all try and do a full SDM on GBIF data
- No lectures, both classes will be for working
- Hand in at the end of next week
- There will be no assignment this week
- This assignment can be used as part of the final project


## Final Projects

- Can be done in groups (max 6) or individually, but everyone will hand in their own work
- Groups can share code / work together on working out how to run the models
- Three Sub-assignments:
  1. Project Plan
  2. Project Code and Results
  3. Project Presentation (Lightning Talk)
- Details will be up on the course website by next week's first class
  

## 1. Project Plan (Preliminary Research)

- A form on Canvas to fill out with:
  - Name of chosen species
  - Group members if applicable
  - Written Sections:
    - *Species:* Background research on the species (including drawing from academic article if available), their natural history.
    - *Variables:* In an ideal world, what variables would be the most useful to model this species, based on what you discover about their biology?
    - *Models:* What kind of model will you run? Research and choose 1 method, briefly explain how it works to the best of your ability. What hyper-parameters does it have?
  - All sections will be short (less than 300 words)


## Next Week

- Do an SDM on your GBIF data


